Accessing biological data with semantic web technologies

### 2013-12-30 article.tex
Some clean up, like proper references

Taking advantage of existing ontologies. The choice is somewhat arbitrary and no ontological
analysis has been done, e.g. on specified domain and ranges of predicates and implied
properties of classes. The following ontologies are used:
SIO is the Semanticscience Integrated Ontology;
TMO is the Translational Medicine Ontology;
%SO is the Sequence Types and Features Ontology;
GO is the Gene Ontology;
BIBO is the Bibliographic Ontology;
CiTO is the Citation Typing Ontology;
... Prefixes used in the term Uniform Resource Identifiers (URIs) include
\textit{sio} for \url{http://semanticscience.org/resource/},
\textit{tmo} for \url{http://www.w3.org/2001/sw/hcls/ns/transmed/},
\textit{obo} for \url{http://purl.obolibrary.org/obo/},
\textit{bibo} for \url{http://purl.org/ontology/bibo/},
\textit{cito} for \url{http://purl.org/spar/cito/} ...

We can define these prefixes in the store with the following code,
the first prefix in the full syntax and the other namespaces
leaving out the parameters names:

@@

\begin{verbatim}
findAnnotations = paste(
  "prefix up: <http://purl.uniprot.org/core/>",
  "prefix ens: <http://purl.uniprot.org/ensembl/>",
  "prefix rdfs: <http://www.w3.org/2000/01/rdf-schema#>",
  "SELECT DISTINCT ?protein ?mnemonic ?comment WHERE {",
  "  ?transcript up:transcribedFrom ens:ENSG00000012048 .",
  "  ?protein a up:Protein ;",
  "    rdfs:seeAlso ?transcript ;",
  "    up:mnemonic ?mnemonic ;",
  "    up:annotation ?annot .",
  "  ?annot rdfs:comment ?comment .",
  "}"
)
annotations = sparql.remote(
  endpoint="http://beta.sparql.uniprot.org/",
  sparql=findAnnotations, jena=FALSE
)
\end{verbatim}

>>>>

Taking advantage of existing ontologies. The choice is somewhat arbitrary and no ontological
analysis has been done, e.g. on specified domain and ranges of predicates and implied
properties of classes. The following ontologies are used:
SIO is the Semanticscience Integrated Ontology;
TMO is the Translational Medicine Ontology~\cite{Luciano2011};
%SO is the Sequence Types and Features Ontology;
GO is the Gene Ontology;
BIBO is the Bibliographic Ontology~\cite{d2009bibliographic}; and, 
CiTO is the Citation Typing Ontology~\cite{Shotton2010}.

Prefixes used in the term Uniform Resource Identifiers (URIs) include
\textit{sio} for \url{http://semanticscience.org/resource/},
\textit{tmo} for \url{http://www.w3.org/2001/sw/hcls/ns/transmed/},
\textit{obo} for \url{http://purl.obolibrary.org/obo/},
\textit{bibo} for \url{http://purl.org/ontology/bibo/}, and
\textit{cito} for \url{http://purl.org/spar/cito/}.
We can add these prefixes to the store. The following code examples
shows the full syntax in the first addition, but leaves out the
parameters names in the following additions:

@@

\begin{footnotesize}
\begin{verbatim}
findAnnotations = paste(
  "prefix up: <http://purl.uniprot.org/core/>",
  "prefix ens: <http://purl.uniprot.org/ensembl/>",
  "prefix rdfs: <http://www.w3.org/2000/01/rdf-schema#>",
  "SELECT DISTINCT ?protein ?mnemonic ?comment WHERE {",
  "  ?transcript up:transcribedFrom ens:ENSG00000012048 .",
  "  ?protein a up:Protein ;",
  "    rdfs:seeAlso ?transcript ;",
  "    up:mnemonic ?mnemonic ;",
  "    up:annotation ?annot .",
  "  ?annot rdfs:comment ?comment .",
  "}"
)
annotations = sparql.remote(
  endpoint="http://beta.sparql.uniprot.org/",
  sparql=findAnnotations, jena=FALSE
)
\end{verbatim}
\end{footnotesize}
